---
title: "browse_results"
author: "Jovan Tanevski"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Setup

Load necessary libraries

```{r setup, message=FALSE}
library(stringr)
library(dplyr)
library(readr)
library(mistyR)
library(ggplot2)
library(future)
library(factoextra)

plan(multisession)
```

# Output collection

Find the location of the results for all samples

```{r}
outputs <- str_subset(list.dirs("output"), "processed") %>%
  str_subset("failed", negate = TRUE) %>%
  str_subset("slide_seq_processed-[0-9]$", negate = TRUE)
```

Collect the results for all samples, single modality and all replicates

```{r}
misty.results.codex <- collect_results(str_subset(outputs, "codex"))
misty.results.merfish <- collect_results(str_subset(outputs, "merfish"))
misty.results.ligrcp <- collect_results(str_subset(outputs, "ligrcp"))
misty.results.ligpath <- collect_results(str_subset(outputs, "ligpath"))
```

# Get metadata

```{r}
sample.meta <- read.delim("data/HTAPP_MBC_spatial_annotations.tsv",
  na.strings = ""
)
```



# Plot basic results

## CODEX

Filter only genes with mean gain in variance explained of .5% or more to plot the gain and view contributions
```{r}
misty.results.codex %>%
  plot_improvement_stats(trim = 0.5) %>%
  plot_view_contributions(trim = 0.5)
```

Plot interaction heatmaps
```{r}
misty.results.codex %>%
  plot_interaction_heatmap("intra", cutoff = 4, clean = TRUE) %>%
  plot_interaction_heatmap("juxta.15", cutoff = 1, clean = TRUE, trim = 0.5) %>%
  plot_interaction_heatmap("para.100", cutoff = 1, clean = TRUE, trim = 0.5)
```


Plot contrasts
```{r}
misty.results.codex %>%
  plot_contrast_heatmap("intra", "juxta.15", cutoff = 1, trim = 0.5) %>%
  plot_contrast_heatmap("intra", "para.100", cutoff = 1, trim = 0.5) %>%
  plot_contrast_heatmap("para.100", "juxta.15", cutoff = 1, trim = 0.5) %>%
  plot_contrast_heatmap("juxta.15", "para.100", cutoff = 1, trim = 0.5)
```

Plot interaction communities 

```{r}
misty.results.codex %>%
  plot_interaction_communities("intra", cutoff = 4) %>%
  plot_interaction_communities("juxta.15", cutoff = 1) %>%
  plot_interaction_communities("para.100", cutoff = 1)
```

Signatures and clustering

```{r}
signature.per.codex <- misty.results.codex %>%
  extract_signature("performance", trim = 0.5) %>%
  mutate(sample = str_extract(sample, "HTAPP(-[:alnum:]+){3}"))

signature.per.pca <- signature.per.codex %>%
  select(-sample) %>%
  prcomp()

signature.per.pca.ann <- left_join(bind_cols(
  signature.per.codex %>% select(sample),
  as.data.frame(signature.per.pca$x)
),
sample.meta,
by = c("sample" = "name")
)


ggplot(signature.per.pca.ann, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = site_biopsy)) +
  theme_classic()

ggplot(signature.per.pca.ann, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = receptors_biopsy)) +
  theme_classic()


fviz_pca_var(signature.per.pca, col.var = "cos2", gradient.cols = c("#666666", "#377EB8", "#E41A1C"), repel = TRUE) + theme_classic()
```


## MERFISH

We combine the information from binned and unbinned samples.

Filter only genes with mean gain in variance explained of 2 or more to plot the gain and view contributions

MERFISH shows much better performance improvement than CODEX, but captures much less information in the intraview on average.

```{r}
misty.results.merfish %>%
  plot_improvement_stats(trim = 2) %>%
  plot_view_contributions(trim = 2)
```

MERFISH shows much better performance improvement than CODEX, but captures much less information in the intraview on average.

```{r}
misty.results.merfish$improvements %>%
  filter(measure == "intra.R2") %>%
  pull(value) %>%
  hist(main = "MERFISH distribution of variance explained in intraview")

misty.results.merfish$improvements %>%
  filter(measure == "gain.R2") %>%
  pull(value) %>%
  hist(main = "MERFISH distribution of gain in variance explained")

misty.results.codex$improvements %>%
  filter(measure == "intra.R2") %>%
  pull(value) %>%
  hist(main = "CODEX distribution of variance explained in intraview")

misty.results.codex$improvements %>%
  filter(measure == "gain.R2") %>%
  pull(value) %>%
  hist(main = "CODEX distribution of gain in variance explained")
```


Plot interaction heatmaps
```{r}
misty.results.merfish %>%
  plot_interaction_heatmap("intra", cutoff = 6, clean = TRUE) %>%
  plot_interaction_heatmap("juxta.15", cutoff = 1.5, clean = TRUE, trim = 2) %>%
  plot_interaction_heatmap("para.100", cutoff = 1.5, clean = TRUE, trim = 2)
```

Plot contrasts
```{r}
misty.results.merfish %>%
  plot_contrast_heatmap("intra", "juxta.15", cutoff = 1.5, trim = 2) %>%
  plot_contrast_heatmap("intra", "para.100", cutoff = 1.5, trim = 2) %>%
  plot_contrast_heatmap("para.100", "juxta.15", cutoff = 1.5, trim = 2) %>%
  plot_contrast_heatmap("juxta.15", "para.100", cutoff = 1.5, trim = 2)
```


Plot interaction communities 

```{r}
misty.results.merfish %>%
  plot_interaction_communities("intra", cutoff = 6) %>%
  plot_interaction_communities("juxta.15", cutoff = 2) %>%
  plot_interaction_communities("para.100", cutoff = 2)
```

Signatures and clustering

We observe better clustering based on biopsy site than receptor status

```{r}
signature.per.merfish <- misty.results.merfish %>%
  extract_signature("performance", trim = 2) %>%
  mutate(sample = str_extract(sample, "HTAPP(-[:alnum:]+){3}"))

signature.per.pca <- signature.per.merfish %>%
  select(-sample) %>%
  prcomp()

signature.per.pca.ann <- left_join(bind_cols(
  signature.per.merfish %>% select(sample),
  as.data.frame(signature.per.pca$x)
),
sample.meta,
by = c("sample" = "name")
)

ggplot(signature.per.pca.ann, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = site_biopsy)) +
  theme_classic()

ggplot(signature.per.pca.ann, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = receptors_biopsy)) +
  theme_classic()

fviz_pca_var(signature.per.pca, col.var = "cos2", gradient.cols = c("#666666", "#377EB8", "#E41A1C"), repel = TRUE) + theme_classic()
```


## Merge CODEX and MERFISH results

Filter only genes with mean gain in variance explained of 1 or more to plot the gain and view contributions
```{r}
misty.results.merged <- collect_results(str_subset(outputs, "(codex|merfish)"))

misty.results.merged %>%
  plot_improvement_stats(trim = 2) %>%
  plot_view_contributions(trim = 2)
```

Plot interaction heatmaps
```{r}
misty.results.merged %>%
  plot_interaction_heatmap("intra", cutoff = 8, clean = TRUE) %>%
  plot_interaction_heatmap("juxta.15", cutoff = 1.5, clean = TRUE, trim = 2) %>%
  plot_interaction_heatmap("para.100", cutoff = 1.5, clean = TRUE, trim = 2) %>%
  plot_contrast_heatmap("intra", "juxta.15", cutoff = 1.5, trim = 2) %>%
  plot_contrast_heatmap("intra", "para.100", cutoff = 1.5, trim = 2)
```


Signatures

There are no common targets with improvements above 1% between the two technologies, therefore we extract the signature using all markers, although we expect to have a technology based batch effect.

```{r}
sig.perf <- extract_signature(misty.results.merged, "performance")
sig.perf.pca <- sig.perf %>%
  select(-sample) %>%
  prcomp()

sig.perf.join <- sig.perf %>%
  mutate(
    tech = str_extract(sample, "(merfish_processed|merfish_bin|codex)"),
    sample = str_extract(sample, "HTAPP-[0-9]{3}-SMP-[0-9]{3,4}")
  ) %>%
  select(tech, sample) %>%
  bind_cols(as.data.frame(sig.perf.pca$x)) %>%
  left_join(sample.meta, by = c("sample" = "name"))

ggplot(sig.perf.join, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = tech)) +
  theme_classic()

ggplot(sig.perf.join, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = site_biopsy)) +
  theme_classic()
```


## SlideSeq

Filter only ligands then pathways with mean gain in variance explained of 1 or more to plot the gain and view contributions

```{r}
misty.results.ligrcp %>%
  plot_improvement_stats(trim = 0.5) %>%
  plot_view_contributions(trim = 0.5)
```

```{r}
misty.results.ligpath %>%
  plot_improvement_stats(trim = 1) %>%
  plot_view_contributions(trim = 1)
```

Plot interaction heatmaps
```{r}
misty.results.ligrcp %>%
  plot_interaction_heatmap("intra", cutoff = 5, clean = TRUE) %>%
  plot_interaction_heatmap("juxta.25", cutoff = 3, clean = TRUE, trim = 0.5) %>%
  plot_interaction_heatmap("para.150", cutoff = 2, clean = TRUE, trim = 0.5)

misty.results.ligpath %>%
  plot_interaction_heatmap("intra", cutoff = 1.5, clean = TRUE) %>%
  plot_interaction_heatmap("juxta.25", cutoff = 3, clean = TRUE, trim = 1) %>%
  plot_interaction_heatmap("para.150", cutoff = 1.5, clean = TRUE, trim = 1)
```

Plot intrinsic pathway communities

```{r}
misty.results.ligpath %>% plot_interaction_communities("intra", cutoff = 1.5)
```

### Cell-type based analysis

We are interested in predicting the probability of a cell being of a cell-type of interest, by looking at the distribution of cell types in the neighborhood of 100 cells.

More conservative trimming of above 10% variance explained since the intraview is bypassed.
The contribution of the intraview can be interpreted as the amount of variance captured only by the mean of the probability estimate for that cell-type. For a good result it is expected to be close to zero.

```{r}
misty.results.ctype <- collect_results(str_subset(outputs, "ctype"))

misty.results.ctype %>%
  plot_improvement_stats(trim = 10) %>%
  plot_view_contributions(trim = 10)
```

Plot neighborhood interactions

```{r}
misty.results.ctype %>%
  plot_interaction_heatmap("juxta.25", trim = 10, cutoff = 0.5) %>%
  plot_interaction_heatmap("para.150", trim = 10, cutoff = 0.5)
```

Plot contrasts

```{r}
misty.results.ctype %>%
  plot_contrast_heatmap("para.150", "juxta.25", trim = 10, cutoff = 0.5) %>%
  plot_contrast_heatmap("juxta.25", "para.150", trim = 10, cutoff = 0.5)
```


Plot communities

```{r}
misty.results.ctype %>%
  plot_interaction_communities("juxta.25", cutoff = 1) %>%
  plot_interaction_communities("para.150", cutoff = 1)
```


```{r eval=FALSE, include=FALSE}
# features <- colnames(sample.meta)[-c(1,2,12)]
# features %>% map_dbl(~silhouette(as.numeric(factor(signature.per.pca.ann  %>% pull(.x))), dist = dist(signature.per.pca$x[,1:2]))[,3] %>% mean())
```
