---
title: "cellcomm_misty"
author: "Jovan Tanevski"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Setup

```{r setup, message  = FALSE}
library(reticulate)
library(mistyR)
library(future)
library(dplyr)
library(purrr)
library(OmnipathR)
library(stringr)
library(progeny)

use_python("/usr/local/bin/python3")
plan(multisession, workers = 5)
```


# Link to python to extract celltype information and geometry

Requires `python3` and anndata installed (`pip install anndata`).

```{python}
import anndata
import numpy

def read_and_extract(datapath):
  adata = anndata.read(datapath)
  ots = adata.obsm["OT"]
  counts = adata.obsm["counts"]
  geom = adata.obs[["x_orig", "y_orig"]]
  markers = adata.uns["counts_var"]
  return ots, counts, geom, markers

```

# Create ligand-receptor oriented mistyR pipeline

Get ligand and receptor symbols from Omnipath

```{r}
extract_symbols <- function(op.source) {
  proteins <- op.source %>%
    filter(entity_type == "protein") %>%
    pull(genesymbol)
  complexes <- op.source %>%
    filter(entity_type == "complex") %>%
    pull(genesymbol) %>%
    str_remove("COMPLEX:") %>%
    str_split("_") %>%
    unlist()
  return(union(proteins, complexes) %>% make.names())
}

op.ligands <- import_omnipath_intercell(
  categories = "ligand",
  secreted = TRUE,
  consensus_percentile = 50
)

ligands <- extract_symbols(op.ligands)

op.receptors <- import_omnipath_intercell(
  categories = "receptor",
  secreted = FALSE,
  consensus_percentile = 50
)

receptors <- extract_symbols(op.receptors)
```


Define and run ligand receptor and ligand pathway oriented mistyR pipelinesin SlideSeq data. Relate expression of genes annotated as ligands in the neighborhood of each cell to its intracellular receptor expression or estimated pathway activities.

```{r}
# 364 is a tricky sample
(list.files("data", ".h5ad", recursive = TRUE, full.names = TRUE) %>%
  keep(~ str_detect(.x, "slide_seq")))[-4] %>%
  walk(function(datapath) {
    data <- py$read_and_extract(datapath)
    counts <- as.matrix(data[[2]])
    transformed <- log((counts / rowSums(counts)) * 1e5 + 1)
    expr <- as.data.frame(transformed) %>%
      `colnames<-`(make.names(data[[4]]))
    pos <- data[[3]]

    unique(str_extract(rownames(pos), "-\\d$")) %>% walk(function(replicate) {
      output.folder <- paste0(
        str_replace(
          str_remove(datapath, ".h5ad"),
          "data", "output"
        ), replicate
      )

      if (!dir.exists(output.folder)) dir.create(output.folder)

      ind <- str_which(rownames(pos), paste0(replicate, "$"))

      if (length(list.files(output.folder)) < 2) {
        ligand.views <- create_initial_view(expr[ind, ]) %>%
          select_markers(
            "intraview",
            where(\(x) sd(x) > 0) &
              any_of(ligands)
          ) %>%
          add_juxtaview(pos[ind, ], neighbor.thr = 25) %>%
          add_paraview(pos[ind, ], l = 150, zoi = 25)
      }


      if (!(dir.exists(paste0(output.folder, "/ligrcp")) |
        dir.exists(paste0(output.folder, "/ligrcp_failed")))) {
        misty.views.ligrcp <- create_initial_view(expr[ind, ]) %>%
          select_markers(
            "intraview",
            where(\(x) length(unique(x)) >= 10) &
              any_of(receptors)
          ) %>%
          add_views(list(
            juxtaview.25 = ligand.views[["juxtaview.25"]],
            paraview.150 = ligand.views[["paraview.150"]]
          ))

        tryCatch(
          run_misty(misty.views.ligrcp, results.folder = paste0(output.folder, "/ligrcp")),
          error = function(e) {
            file.rename(
              paste0(output.folder, "/ligrcp"),
              paste0(output.folder, "/ligrcp_failed")
            )
          }
        )
      }
      if (!(dir.exists(paste0(output.folder, "/ligpath")) |
        dir.exists(paste0(output.folder, "/ligpath_failed")))) {
        path.activity <- progeny(expr = t(expr[ind, ]) %>% `rownames<-`(data[[4]]), 
                                 top = 500, scale = TRUE) %>%
          data.frame()

        misty.views.ligpath <- create_initial_view(path.activity) %>%
          select_markers(
            "intraview",
            where(\(x) length(unique(x)) >= 10)
          ) %>%
          add_views(list(
            juxtaview.25 = ligand.views[["juxtaview.25"]],
            paraview.150 = ligand.views[["paraview.150"]]
          ))

        tryCatch(
          run_misty(misty.views.ligpath, results.folder = paste0(output.folder, "/ligpath")),
          error = function(e) {
            file.rename(
              paste0(output.folder, "/ligpath"),
              paste0(output.folder, "/ligpath_failed")
            )
          }
        )
      }
    })
  })
```
